---
title: "(Phospho-)flow cytometry gating strategy"
author: "mwitmond"
date: "2024-08-06"
output:
  workflowr::wflow_html:
    toc: TRUE
    toc_float: 
      collapsed: FALSE
    toc_depth: 2
    code_folding: hide
editor_options:
  chunk_output_type: console
---


## Set-up

```{r setup, message=F, warning=F}
knitr::opts_chunk$set(
  message = F, warning = F, echo = T, eval = T
)

# Load required packages
source("code/packages_FACS.R", local = knitr::knit_global())
```

```{r fig_settings}
row_order = c("A", "B", "C", "D", "E", "F", "G", "H")
col_order = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
panel_labels <- c("a", "b", "c","d", "e", "f", "g", "h", "i", "j", "k", "l", "m")

textsize <- theme(axis.text.x = element_text(colour = "grey", size = 11), #, face = "bold"
                  axis.text.y = element_text(colour = "grey", size = 11),
                  axis.title = element_text(colour = "black", size = 12), 
                  legend.title = element_text(colour = "black", size = 12),
                  # legend.title = element_blank(), 
                  legend.text = element_text(colour = "grey", size = 11), 
                  strip.text.x = element_text(colour = "black", size = 12)
)

textsize_small <- theme(text = element_text(size = 7, family = "sans", colour = "black"),
                        plot.title = element_text(size = 8)
)
textsize_medium <- theme(text = element_text(size = 10, family = "sans", colour = "black"),
                         strip.text.x = element_text(size = 10), 
                         plot.title = element_text(size = 10)
)

colors_dark9 <- c("#4daf4a", "#984ea3", "#377eb8", "#ff7f00", "#f781bf", "#ffff33", "#e41a1c", "#a65628", "#999999")
colors_light12 <- c("#8dd3c7","#ffffb3","#bebada","#fb8072","#80b1d3","#fdb462","#b3de69","#fccde5","#d9d9d9","#bc80bd","#ccebc5","#ffed6f")
colors_paired10 <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fdbf6f", "#ff7f00", "#cab2d6", "#6a3d9a", "#fb9a99", "#e31a1c")

colors_blue9 <- c("#f7fbff", "#deebf7", "#c6dbef", "#9ecae1", "#6baed6", "#4292c6", "#2171b5", "#08519c", "#08306b")
colors_green9 <- c("#f7fcf5", "#e5f5e0", "#c7e9c0", "#a1d99b", "#74c476", "#41ab5d", "#238b45", "#006d2c", "#00441b")
colors_purple9 <- c("#fcfbfd", "#efedf5", "#dadaeb", "#bcbddc", "#9e9ac8", "#807dba", "#6a51a3", "#54278f", "#3f007d")
colors_red9 <- c("#fff5f0", "#fee0d2", "#fcbba1", "#fc9272", "#fb6a4a", "#ef3b2c", "#cb181d", "#a50f15", "#67000d")
colors_orange9 <- c("#fff5eb", "#fee6ce", "#fdd0a2", "#fdae6b", "#fd8d3c", "#f16913", "#d94801", "#a63603", "#7f2704")
colors_grey9 <- c("#ffffff", "#f0f0f0", "#d9d9d9", "#bdbdbd", "#969696", "#737373", "#525252", "#252525", "#000000")

colors_yb9 <- c("#ffffd9", "#edf8b1", "#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494", "#081d58")
colors_pr9 <- c("#f7f4f9", "#e7e1ef", "#d4b9da", "#c994c7", "#df65b0", "#e7298a", "#ce1256", "#980043", "#67001f")

colors_proteins <- c("pCD79a (Y182)" = "#892BE1", "pSYK (Y525/Y526)" = "#1E8FFF", "pPLCy2 (Y759)" = "#20B1A9")
colors_stim <- c("PBS" = "#737373", "anti-Ig" = "#74c476", "H2O2" = "#6EA6CD")
colors_stim_clean <- c("Basal" = "#737373", "Activated (anti-Ig)" = "#74c476", "Activated (H2O2)" = "#6EA6CD")
colors_cell <- c("HBL1" = "#df65b0", "DAUDI" = "#41b6c4")
```



## DS091: ImmunostainingBio

### Data

Read the .fcs files and create a FlowSet object with metadata
```{r data_DS091}
# Read files with the FlowSet function
fs_A_DS091 <- read.flowSet(path = "data/DS091_ImmunostainingBio/flow/fcs_plateA", pattern = ".fcs", alter.names = T)
```

```{r metadata_DS091}
# Extract information on wellID, row and column from the file name
# Format of file name: wellID_number_date_time.fcs
# Plate A
pData(fs_A_DS091)$tubeID <- gsub("(.*)_.*_.*_.*.fcs", "\\1", sampleNames(fs_A_DS091))
pData(fs_A_DS091)$rowID <- gsub("(.).*", "\\1", pData(fs_A_DS091)$tubeID)
pData(fs_A_DS091)$colID <- as.numeric(gsub(".(.*)", "\\1", pData(fs_A_DS091)$tubeID))
pData(fs_A_DS091)$expID <- "DS091"
pData(fs_A_DS091)$plateID <- "A"

kable(pData(fs_A_DS091) %>% head(5)) %>% kable_styling(bootstrap_options = c("striped", "hover"))

# Create a GatingSet object from the FlowSet object
gs_DS091 <- GatingSet(fs_A_DS091)
```


### Debris

Define the gate on debris (SSC-A vs FSC-A): Exclude the lower left data points and those data point that fall at/off the upper and right edges of the plot
```{r debris_DS091, fig.width=5, fig.height=4}
# # Create a GatingSet object from the FlowSet object
# gs_DS091 <- GatingSet(fs_HBL1)

# Manually adjust the numbers and check in the graph to set the right gate
g_debris <- polygonGate(filterId = "Debris", "FSC.A" = c(3e4, 26e4, 26e4, 1.5e5), "SSC.A" = c(1e4, 1e4, 26e4, 26e4)) # define gate

ggcyto(gs_DS091[[1]], aes(x = FSC.A, y = SSC.A), subset = "root") +
  geom_hex(bins = 200) +
  geom_gate(g_debris) +
  ggcyto_par_set(limits = "instrument") +
  labs(x = "Cell size (FSC-A)", y = "Cell complexity (SSC-A)") +
  theme_bw() +
  textsize
```
***Figure:** Cell size vs cell complexity and debris gate. One exemplary sample.*  

```{r fig_debris_DS091, fig.width=12, fig.height=2}
# Apply the gate to all data
gs_pop_add(gs_DS091, g_debris)
recompute(gs_DS091)

# Plate A
ggcyto(subset(gs_DS091, plateID == "A"), aes(x = FSC.A, y = SSC.A), subset = "root") +
  geom_hex(bins = 100) +
  geom_gate("Debris") +
  geom_stats(adjust = 0.8) +
  ggcyto_par_set(limits = "instrument" ) +
  facet_grid(vars(rowID), vars(factor(colID, levels = col_order))) +
  labs(x = "Cell size (FSC-A)", y = "Cell complexity (SSC-A)", title = "Non-debris from root (HBL1)") +
  theme_bw() +
  theme(legend.position = "none") + 
  textsize_small
```

```{r debris_DS091_fig, fig.width=3, fig.height=3}
# Figure for gating strategy
fig_debris_DS091 <- ggcyto(gs_DS091[[1]], aes(x = FSC.A, y = SSC.A), subset = "root") +
  geom_hex(bins = 200) +
  geom_gate("Debris") +
  geom_stats(adjust = c(0.8, 0.1)) +
  ggcyto_par_set(limits = "instrument") +
  facet_wrap(vars(factor(tubeID, label = "Tube 1"))) +
  labs(x = "Cell size (FSC-A)", y = "Cell complexity (SSC-A)", title = "Non-debris gate") +
  theme_bw() +
  theme(legend.position = "none") +
  textsize_medium
fig_debris_DS091

# Save figure as png
ggsave(
  fig_debris_DS091,
  filename = "output/figures/gating_figs/fig_debris_DS091.png",
  width = 3,
  height = 3,
  units = "in",
  dpi = 300
  )
```


### Singlets

Define the gate on singlets (FSC-W vs FSC-H): Exclude all data points that do not fall in the mass of data points
```{r singlets_DS091, fig.width=5, fig.height=4}
# Manually adjust the numbers and check in the graph to set right gate
g_singlets <- polygonGate(filterId = "Singlets", "FSC.H" = c(3e4, 6.5e4, 7.5e4, 2.5e4), "FSC.W" = c(7.7e4, 9e4, 1.4e5, 1.4e5)) # define gate

ggcyto(gs_DS091[[1]], aes(x = FSC.H, y = FSC.W), subset = "Debris") +
  geom_hex(bins = 200) +
  geom_gate(g_singlets) +
  ggcyto_par_set(limits = "instrument") +
  labs(x = "Cell size height (FSC-H)", y = "Cell size width (FSC-W)") +
  theme_bw() +
  textsize
```
***Figure:** Cell size height vs cell size width and singlet cell gate. One exemplary sample.* 

```{r fig_singlets_DS091, fig.width=12, fig.height=2}
# Apply the gate to all data
gs_pop_add(gs_DS091, g_singlets, parent = "Debris")
recompute(gs_DS091) 

# Plate A
ggcyto(subset(gs_DS091, plateID == "A"), aes(x = FSC.H, y = FSC.W), subset = "Debris") +
  geom_hex(bins = 100) +
  geom_gate("Singlets") +
  geom_stats(adjust = c(1.5, 0)) +
  ggcyto_par_set(limits = "instrument") +
  facet_grid(vars(rowID), vars(factor(colID, levels = col_order))) +
  labs(x = "Cell size height (FSC-H)", y = "Cell size width (FSC-W)", title = "Singlets from non-debris (HBL1)") +
  theme_bw() +
  theme(legend.position = "none") + 
  textsize_small
```

```{r singlets_DS091_fig, fig.width=3, fig.height=3}
# Figure for gating strategy
fig_singlets_DS091 <- ggcyto(gs_DS091[[1]], aes(x = FSC.H, y = FSC.W), subset = "Debris") +
  geom_hex(bins = 150) +
  geom_gate("Singlets") +
  geom_stats(adjust = c(1.4, 0.1)) +
  ggcyto_par_set(limits = "instrument") +
  facet_wrap(vars(factor(tubeID, label = "Tube 1"))) +
  labs(x = "Cell size height (FSC-H)", y = "Cell size width (FSC-W)", title = "Singlets gate") +
  theme_bw() +
  theme(legend.position = "none") +
  textsize_medium
fig_singlets_DS091

# Save figure as png
ggsave(
  fig_singlets_DS091,
  filename = "output/figures/gating_figs/fig_singlets_DS091.png",
  width = 3,
  height = 3,
  units = "in",
  dpi = 300
  )
```


### Live cells (fixed)

Define the gate on live/dead (fixed) cells (BV421 vs FSC-A): Exclude the upper data points and those data point that fall at/off the right edges of the plot
```{r live_DS091, fig.width=5, fig.height=4}
# Manually adjust the numbers and check in the graph to set right gate
g_live <- polygonGate(filterId = "Live", "FSC.A" = c(3e4, 1.6e5, 1.6e5, 3e4), "BV421.A" = c(1e2, 1e2, 2e4, 1.5e3)) # define gate

ggcyto(gs_DS091[[1]], aes(x = FSC.A, y = BV421.A), subset = "Singlets") +
  geom_hex(bins = 200) +
  geom_gate(g_live) +
  scale_x_log10() +
  scale_y_log10() +
  coord_cartesian(ylim = c(1e2, 2e5)) +
  labs(x = "Cell size (FSC-A)", y = "Live/death stain (BV421-A)") +
  theme_bw() +
  textsize
```
***Figure:** Cell size vs live/death stain and live cell gate. One exemplary sample.* 

```{r fig_live_DS091, fig.width=12, fig.height=2}
# Apply the gate to all data
gs_pop_add(gs_DS091, g_live, parent = "Singlets")
recompute(gs_DS091) 

# Plate A
ggcyto(subset(gs_DS091, plateID == "A"), aes(x = FSC.A, y = BV421.A), subset = "Singlets") +
  geom_hex(bins = 100) +
  geom_gate("Live") +
  geom_stats(adjust = 0.8) +
  scale_x_log10() +
  scale_y_log10() +
  # coord_cartesian(ylim = c(1e2, 2e5)) +
  facet_grid(vars(rowID), vars(factor(colID, levels = col_order))) +
  labs(x = "Cell size (FSC-A)", y = "Live/death stain (BV421-A)", title = "Live cells from singlets (HBL1)") +
  theme_bw() +
  theme(legend.position = "none") + 
  textsize_small
```

```{r live_DS091_fig, fig.width=3, fig.height=3}
# Figure for gating strategy
fig_live_DS091 <- ggcyto(gs_DS091[[1]], aes(x = FSC.A, y = BV421.A), subset = "Singlets") +
  geom_hex(bins = 150) +
  geom_gate("Live") +
  geom_stats(adjust = c(0.75, 0)) +
  scale_x_log10() +
  scale_y_log10(limits = c(1e2, 2e5)) +
  # coord_cartesian(ylim = c(1e2, 2e5)) +
  facet_wrap(vars(factor(tubeID, label = "Tube 1"))) +
  labs(x = "Cell size (FSC-A)", y = "Live/death stain (BV421-A)", title = "Live cell gate") +
  theme_bw() +
  theme(legend.position = "none") +
  textsize_medium
fig_live_DS091

# Save figure as png
ggsave(
  fig_live_DS091,
  filename = "output/figures/gating_figs/fig_live_DS091.png",
  width = 3,
  height = 3,
  units = "in",
  dpi = 300
  )
```


### Summary

```{r gating_DS091}
# Remove previous file if it exist
if(file.exists(paste0("output/DS091_ImmunostainingBio/", "flow_gated"))) {
  unlink(paste0("output/DS091_ImmunostainingBio/", "flow_gated"), recursive = TRUE)
}

# Save dataset
save_gs(gs_DS091, path = paste0("output/DS091_ImmunostainingBio/", "flow_gated"))

gc()

# Gating strategy
plot(gs_DS091, bool = TRUE)
```
***Figure:** Gating strategy.* 



### Population statistics

```{r cell_counts_DS091}
# Get cell counts
counts_DS091 <- gs_pop_get_count_fast(gs_DS091)
counts_DS091 <- merge(counts_DS091, pData(fs_A_DS091), by = "name")
counts_DS091$PercentParent <- (counts_DS091$Count / counts_DS091$ParentCount) * 100

# Calculate percentage from root
counts_root_DS091 <- counts_DS091 %>% dplyr::filter(Parent == "root") %>% dplyr::select(name, RootCount = ParentCount)
counts_DS091 <- left_join(counts_DS091, counts_root_DS091)
counts_DS091$PercentRoot <- (counts_DS091$Count / counts_DS091$RootCount) * 100

# Save the number of cells per well per gate as .csv file
counts_save_DS091 <- counts_DS091[ , c("plateID", "tubeID", "name", "Population", "Count", "ParentCount", "PercentParent", "RootCount", "PercentRoot")]
write_csv(counts_save_DS091, file = "output/DS091_ImmunostainingBio/flow_ann/gated_counts_DS091.csv")
```

```{r fig_cells_DS091, fig.width=6, fig.height=4}
# Plate A
counts_final_A_DS091 <- subset(counts_DS091, Population == "/Debris/Singlets/Live" & plateID == "A")
raw_map(data = counts_final_A_DS091$Count, well = counts_final_A_DS091$tubeID, plate = 96, size = 10) +
  scale_fill_viridis() +
  labs(title = "Single cell counts") +
  theme_bw() +
  textsize
```
***Figure:** Number of live single cells after gating (counts per sample/well).* 

```{r live_cells_DS091}
percent_live_DS091 <- subset(counts_DS091, Population == "/Debris/Singlets/Live")[, c("name", "plateID", "tubeID", "PercentRoot", "PercentParent")]

# Add metadata per well from custom .csv file
metadata_DS091 <- read_csv("data/DS091_ImmunostainingBio/flow/metadata_tubes_DS091.csv")
percent_live_DS091 <- left_join(percent_live_DS091, metadata_DS091)
```

```{r fig_live_percent_DS091, fig.width=4, fig.height=4}
stim_DS091 <- c("Basal", "Activated (anti-Ig)", "Activated (H2O2)")
cell_stim_DS091 <- c("DAUDI - Basal", "DAUDI - Activated (anti-Ig)", "HBL1 - Basal", "HBL1 - Activated (H2O2)")
cell_stim_label_DS091 <- c("DAUDI\nBasal", "DAUDI\nActivated (anti-Ig)", "HBL1\nBasal", "HBL1\nActivated (H2O2)")

ggplot(percent_live_DS091, aes(x = factor(description, levels = cell_stim_DS091, labels = cell_stim_label_DS091), 
                               y = PercentRoot)) +
  geom_point(aes(color = stim_clean, shape = cell_line), size = 2) +
  # facet_wrap(~cell_line, ncol = 1) +
  ylim(0, 100) +
  scale_color_manual(values = colors_stim_clean, breaks = stim_DS091) +
  labs(y = "Percentage live single \ncells from root (%)", x = "") +
  theme_bw() +
  theme(legend.position = "right", legend.justification = "right") +
  textsize_small

ggplot(percent_live_DS091, aes(x = factor(description, levels = cell_stim_DS091, labels = cell_stim_label_DS091), 
                               y = PercentParent)) +
  geom_point(aes(color = stim_clean, shape = cell_line), size = 2) +
  # facet_wrap(~cell_line, ncol = 1) +
  ylim(0, 100) +
  scale_color_manual(values = colors_stim_clean, breaks = stim_DS091) +
  labs(y = "Percentage live \ncells from singlets (%)", x = "") +
  theme_bw() +
  theme(legend.position = "right", legend.justification = "right") +
  textsize_small
```
***Figure:** Percentage of live single cells from root/parent.*

```{r clear_DS091}
# Remove unnecessary files to clear up memory
rm(list = ls(pattern = "_DS091"))
rm(list = ls(pattern = "g_"))
rm(list = ls(pattern = "gs_"))
rm(list = ls(pattern = "counts_"))
gc()
```
