---
title: "ID-seq annotation"
author: "mwitmond"
date: "2024-09-24"
output:
  workflowr::wflow_html:
    toc: TRUE
    toc_float: 
      collapsed: FALSE
    toc_depth: 3
editor_options:
  chunk_output_type: console
---


## Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = F, warning = F, echo = T, eval = T
)

# Load required packages
source("code/packages_seq.R", local = knitr::knit_global())
```

```{r fig_settings}
row_order = c("A", "B", "C", "D", "E", "F", "G", "H")
col_order = c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12")
panel_labels <- c("a", "b", "c","d", "e", "f", "g", "h", "i", "j", "k", "l", "m")

textsize <- theme(axis.text.x = element_text(colour = "grey", size = 11), #, face = "bold"
                  axis.text.y = element_text(colour = "grey", size = 11),
                  axis.title = element_text(colour = "black", size = 12), 
                  legend.title = element_text(colour = "black", size = 12),
                  # legend.title = element_blank(), 
                  legend.text = element_text(colour = "grey", size = 11), 
                  strip.text.x = element_text(colour = "black", size = 12)
)

textsize_small <- theme(text = element_text(size = 7, family = "sans", colour = "black"),
                        plot.title = element_text(size = 8)
)
textsize_medium <- theme(text = element_text(size = 10, family = "sans", colour = "black"),
                         strip.text.x = element_text(size = 10), 
                         plot.title = element_text(size = 10)
)

colors_dark9 <- c("#4daf4a", "#984ea3", "#377eb8", "#ff7f00", "#f781bf", "#ffff33", "#e41a1c", "#a65628", "#999999")
colors_light12 <- c("#8dd3c7","#ffffb3","#bebada","#fb8072","#80b1d3","#fdb462","#b3de69","#fccde5","#d9d9d9","#bc80bd","#ccebc5","#ffed6f")
colors_paired10 <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fdbf6f", "#ff7f00", "#cab2d6", "#6a3d9a", "#fb9a99", "#e31a1c")

colors_blue9 <- c("#f7fbff", "#deebf7", "#c6dbef", "#9ecae1", "#6baed6", "#4292c6", "#2171b5", "#08519c", "#08306b")
colors_green9 <- c("#f7fcf5", "#e5f5e0", "#c7e9c0", "#a1d99b", "#74c476", "#41ab5d", "#238b45", "#006d2c", "#00441b")
colors_purple9 <- c("#fcfbfd", "#efedf5", "#dadaeb", "#bcbddc", "#9e9ac8", "#807dba", "#6a51a3", "#54278f", "#3f007d")
colors_red9 <- c("#fff5f0", "#fee0d2", "#fcbba1", "#fc9272", "#fb6a4a", "#ef3b2c", "#cb181d", "#a50f15", "#67000d")
colors_orange9 <- c("#fff5eb", "#fee6ce", "#fdd0a2", "#fdae6b", "#fd8d3c", "#f16913", "#d94801", "#a63603", "#7f2704")
colors_grey9 <- c("#ffffff", "#f0f0f0", "#d9d9d9", "#bdbdbd", "#969696", "#737373", "#525252", "#252525", "#000000")

colors_yb9 <- c("#ffffd9", "#edf8b1", "#c7e9b4", "#7fcdbb", "#41b6c4", "#1d91c0", "#225ea8", "#253494", "#081d58")
colors_pr9 <- c("#f7f4f9", "#e7e1ef", "#d4b9da", "#c994c7", "#df65b0", "#e7298a", "#ce1256", "#980043", "#67001f")

colors_proteins <- c("pCD79a (Y182)" = "#892BE1", "pSYK (Y525/Y526)" = "#1E8FFF", "pPLCy2 (Y759)" = "#20B1A9")
colors_stim <- c("PBS" = "#737373", "anti-Ig" = "#74c476", "H2O2" = "#6EA6CD")
colors_stim_clean <- c("Basal" = "#737373", "Activated (anti-Ig)" = "#74c476", "Activated (H2O2)" = "#6EA6CD")
colors_cell <- c("HBL1" = "#df65b0", "DAUDI" = "#41b6c4")
```



## DS075: BarcodeCheck

### Raw data

```{r data_DS075}
# Load sparse matrix format all data
counts_PlateA <- read_count_output("data/DS075_BarcodeCheck/counts/PlateA/featurecounts", name = "featurecounts")
counts_PlateB <- read_count_output("data/DS075_BarcodeCheck/counts/PlateB/featurecounts", name = "featurecounts")

meta_wellID <- read_csv("data/DS075_BarcodeCheck/config/DynSign.075_wellBC_metadata_seq.csv")
# meta_wellID <- read_csv("data/DS075_BarcodeCheck/config/meta_antibody_barcodes_samples_new.csv")
meta_Abs_nr <- read_excel("data/DS075_BarcodeCheck/config/meta_featurebarcodes.xlsx")

# Reshape counts in data table format
# Plate A
counts_dtbl_plateA <- data.frame(counts_PlateA) %>% 
  mutate(barcode_name = rownames(counts_PlateA)) %>% #barcode_name #Barcodename
  dplyr::select(barcode_name, everything()) %>%
  gather("well_BC_seq", "counts", 2:c(ncol(counts_PlateA)+1)) %>%
  dplyr::filter(counts > 1) %>% # Remove undetected proteins counts
  left_join(meta_Abs_nr) %>%
  unique() %>%
  # dplyr::filter(!is.na(target_nospace)) %>% # Remove barcodes not in Ab metadata (not conjugated/added in panel)
  left_join(subset(meta_wellID, plate == "A")) %>% # Add sample metadata
  dplyr::filter(!is.na(experiment)) %>%
  arrange(plate, well)

# Plate B
counts_dtbl_plateB <- data.frame(counts_PlateB) %>% 
  mutate(barcode_name = rownames(counts_PlateB)) %>% #barcode_name #Barcodename
  dplyr::select(barcode_name, everything()) %>%
  gather("well_BC_seq", "counts", 2:c(ncol(counts_PlateB)+1)) %>%
  dplyr::filter(counts > 1) %>% # Remove undetected proteins counts
  left_join(meta_Abs_nr) %>%
  unique() %>%
  # dplyr::filter(!is.na(target_nospace)) %>% # Remove barcodes not in Ab metadata (not conjugated/added in panel)
  left_join(subset(meta_wellID, plate == "B")) %>% # Add sample metadata
  dplyr::filter(!is.na(experiment)) %>%
  arrange(plate, well)

counts_dtbl <- full_join(counts_dtbl_plateA, counts_dtbl_plateB)

# Calculate per-sample properties
data_properties <- counts_dtbl %>%
  unique()%>%
  group_by(plate, well_BC_seq) %>%
  summarize(nCount = sum(counts), 
            nProt = n()) %>%
  left_join(meta_wellID) %>% # Add sample metadata 
  arrange(plate, well)
```

***Table**: Unfiltered data-table properties loaded into R*

|                           | nCols (#samples)        | nRows (\# barcodes)     |
|------------------|---------------------------|---------------------------|
| Plate A         | `r ncol(counts_PlateA)` | `r nrow(counts_PlateA)` |

```{r fig_count_prot_DS075, fig.width=8, fig.height=5}
options(repr.plot.width = 6, repr.plot.height = 4)

ggplot(data_properties, aes(log10(nCount), nProt)) +
  geom_point(alpha = 1, size = 1) +
  facet_grid(~plate) +
  # scale_color_manual(values = colors_dark9) +
  theme_bw() +
  labs(title = "Correlation total counts and number of detected barcodes", 
       x = expression("Total counts Log"[10]),
       y = "Detected number of protein barcodes") +
  textsize_small
```

***Figure:** Correlation between total counts and number of detected protein barcodes. We observe many more detected proteins than present in the staining.*

As we have a 96-wells plate, we can also have a look at total counts per plate detected:

```{r fig_plate_totalUMI_DS075, fig.width=8, fig.height=4}
options(repr.plot.width = 10, repr.plot.height = 6)

rng <- range(log10(data_properties$nCount))

plot <- raw_grid(data = log10(data_properties$nCount), well = data_properties$well, plate_id = data_properties$plate, plate = 96, size = 6) +
  scale_fill_viridis(expression("Total counts Log"[10]), limits = c(rng[1], rng[2])) +
  theme_bw()

print(plot)

```

***Figure:** Plate overviews of total counts.*


### Data wrangling

Calculate the following:

-   Percentage contamination

```{r percent_DS075}
# Calculate percentage contamination for each BC
counts_dtbl_percent <- counts_dtbl %>% 
  left_join(data_properties) %>%
  dplyr::rename(counts_BC = counts, 
                counts_total = nCount, 
                measured_BC = barcode_nr_only, 
                measured_BC_long = barcode_name) %>%
  mutate(counts_percent = (counts_BC / counts_total) * 100, 
         plate_well = paste0(plate, "_", well))

# Calculate total percentage contamination
data_total_cont <- counts_dtbl_percent %>%
  dplyr::filter(measured_BC_long != expected_BC_long) %>%
  group_by(plate, well) %>%
  summarize(counts_total_cont = sum(counts_BC), 
            counts_total_percent_cont = sum(counts_percent))
counts_dtbl_percent <- counts_dtbl_percent %>%
  left_join(data_total_cont)
```



### Save data

Finally, we will save the datasets as .csv file

-   IDseq_data_sample.csv: Dataset of counts per protein per sample 

```{r save_DS075}
# Dataset of counts per barcode per well
meta_cols_well_DS075 <- c("plate_well", "plate", "row", "column", "well", 
                          "experiment", "expected_BC", "expected_BC_long", "measured_BC", "measured_BC_long", 
                          "counts_BC", "counts_total", "counts_percent", "counts_total_cont", "counts_total_percent_cont")
data_well_all <- counts_dtbl_percent %>% 
  arrange(plate_well, measured_BC) %>%
  dplyr::select(meta_cols_well_DS075) # reorder columns for a more logical order

write.csv(data_well_all, file = "output/DS075_BarcodeCheck/IDseq_ann/IDseq_data_sample.csv", row.names = F)
```

```{r clear_DS075}
rm(list = ls(pattern = "DS075"))
rm(list = ls(pattern = "Plate"))
gc()
```



## DS078: Blocking1

### Raw data

```{r data_DS078}
# Load sparse matrix format all data
counts_PlateA <- read_count_output("data/DS078_Blocking1/counts/PlateA/featurecounts", name = "featurecounts")
meta_wellID <- read_csv("data/DS078_Blocking1/config/DynSign.078_wellBC_metadata_seq.csv")

meta_Abs_OG <- read_excel("data/DS078_Blocking1/config/BulkIDseq_Ab_info_seq_analysis_202405.xlsx")
meta_Abs_DS078 <- read_csv("data/DS078_Blocking1/config/DynSign.078_Ab_metadata.csv")

# Reshape counts in data table format
counts_dtbl <- data.frame(counts_PlateA) %>% 
  mutate(barcode_name = rownames(counts_PlateA)) %>% #barcode_name #Barcodename
  dplyr::select(barcode_name, everything()) %>%
  gather("well_BC_seq", "counts", 2:c(ncol(counts_PlateA)+1)) %>%
  dplyr::filter(counts >= 1) %>% # Remove undetected proteins counts
  left_join(meta_Abs_DS078) %>%
  unique() %>%
  # dplyr::filter(!is.na(target_nospace)) %>% # Remove barcodes not in Ab metadata (not conjugated/added in panel)
  left_join(subset(meta_wellID, plate == "A")) %>% # Add sample metadata
  dplyr::filter(!is.na(experiment)) %>%
  arrange(plate, well)

counts_dtbl <- counts_dtbl %>%
  mutate(plate_well = paste0(plate, "_", well))

# Calculate per-sample properties
data_properties <- counts_dtbl %>%
  unique()%>%
  group_by(plate, well_BC_seq) %>%
  summarize(nCount = sum(counts), 
            nProt = n()) %>%
  left_join(meta_wellID) # Add sample metadata 
```

***Table**: Unfiltered data-table properties loaded into R*

|                           | nCols (#samples)        | nRows (\# barcodes)     |
|------------------|---------------------------|---------------------------|
| Plate A         | `r ncol(counts_PlateA)` | `r nrow(counts_PlateA)` |

```{r fig_count_prot_DS078, fig.width=8, fig.height=5}
options(repr.plot.width = 6, repr.plot.height = 4)

ggplot(data_properties, aes(log10(nCount), nProt, color = cell_treat)) +
  geom_point(alpha = 1, size = 1) +
  facet_grid(~block_buf) +
  # scale_color_manual(values = colors_dark9) +
  theme_bw() +
  labs(title = "Correlation total counts and number of detected barcodes", 
       x = expression("Total counts Log"[10]),
       y = "Detected number of protein barcodes") +
  textsize_small
```

***Figure:** Correlation between total counts and number of detected protein barcodes. We observe many more detected proteins than present in the staining.*

As we have a 96-wells plate, we can also have a look at total counts per plate detected:

```{r fig_plate_totalUMI_DS078, fig.width=8, fig.height=4}
options(repr.plot.width = 10, repr.plot.height = 6)

rng <- range(log10(data_properties$nCount))

plot <- raw_grid(data = log10(data_properties$nCount), well = data_properties$well, plate_id = data_properties$plate, plate = 96, size = 6) +
  scale_fill_viridis(expression("Total counts Log"[10]), limits = c(rng[1], rng[2])) +
  theme_bw()

print(plot)

```

***Figure:** Plate overviews of total counts.*


### Barcode combination 

Several barcodes are contaminated with other barcodes (all barcodes were sequenced individually to determine the contamination). We combined the counts of barcodes that are attached to the same antibody (same target).

```{r comp_apply_DS078}
counts_comp <- counts_dtbl %>%
  drop_na(target_assign) %>%
  dplyr::select(plate_well, target_nospace, target_present, target_assign, counts) %>%
  group_by(plate_well, target_assign) %>%
  summarise(counts_target = sum(counts))

counts_dtbl <- left_join(counts_dtbl, counts_comp) %>%
  drop_na(target_assign)

# Calculate per-sample properties
data_properties_true <- counts_dtbl %>%
  unique()%>%
  group_by(plate, well_BC_seq) %>%
  summarize(nCount = sum(counts_target, na.rm = TRUE), 
            nProt = n()) %>%
  left_join(meta_wellID) # Add sample metadata 
```

```{r fig_count_raw_comp_DS078, fig.width=8, fig.height=5}
options(repr.plot.width = 6, repr.plot.height = 4)

ggplot(data_properties_true, aes(log10(nCount), nProt, color = cell_treat)) +
  geom_point(alpha = 1, size = 1) +
  facet_grid(~block_buf) +
  # scale_color_manual(values = colors_dark9) +
  theme_bw() +
  labs(title = "Correlation total counts and number of detected barcodes", 
       x = expression("Total counts Log"[10]),
       y = "Detected number of protein barcodes") +
  textsize_small

ggplot(drop_na(counts_dtbl, target_assign), aes(log10(counts), log10(counts_target), color = cell_treat)) +
  geom_point(alpha = 1, size = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  facet_grid(~block_buf) +
  # scale_color_manual(values = colors_conc, name = "iSYK + iBTK + iNFkB") +
  theme_bw() +
  labs(title = "Correlation original counts and combined counts",
       x = expression("Original counts Log"[10]),
       y = expression("Combined counts Log"[10])) +
  textsize_small
```

***Figure:** Correlation between original counts and combined counts, and between compensated counts and detected number of barcodes.*


### Count normalisation

To remove variation between replicates, the data is normalised per target protein with a geometric mean scaling factor: First, the geometric mean of each target is calculated separately for each cell line, and then the median of the protein scaling factors is calculated separately for each sample.

```{r normalisation_DS078}
scaling_factors <- counts_dtbl %>%
  dplyr::filter(staining == "yes") %>%
  # dplyr::filter() %>% #target_nospace == "GAPDH" # modification == "no"
  ungroup() %>%
  group_by(target_nospace) %>%
  mutate(scaling_factor_geo = counts / exp(mean(log((counts)))), 
         true_scaling_factor_geo = counts_target / exp(mean(log((counts_target))))) %>%
  ungroup() %>%
  group_by(description_rep) %>%
  summarize(scaling_factor_geo = median(scaling_factor_geo), 
            true_scaling_factor_geo = median(true_scaling_factor_geo, na.rm = TRUE))

counts_dtbl_norm <- counts_dtbl %>%
  dplyr::filter(staining == "yes") %>%
  left_join(scaling_factors) %>%
  ungroup() %>%
  mutate(counts_norm_geo = counts / scaling_factor_geo, 
         counts_target_norm_geo = counts_target / true_scaling_factor_geo)

counts_dtbl_no <- counts_dtbl %>%
  dplyr::filter(staining == "no")
counts_dtbl_norm <- full_join(counts_dtbl_norm, counts_dtbl_no) %>%
  dplyr::filter(!is.na(target_nospace))

# Calculate per-sample properties
data_properties_norm <- counts_dtbl_norm %>%
  unique()%>%
  group_by(plate, well_BC_seq) %>%
  summarize(nCount_norm = sum(counts_norm_geo), 
            nProt = n()) %>%
  left_join(meta_wellID)

data_properties_norm_true <- counts_dtbl_norm %>%
  unique()%>%
  group_by(plate, well_BC_seq) %>%
  summarize(nCount_norm = sum(counts_target_norm_geo), 
            nProt = n()) %>%
  left_join(meta_wellID)
```

```{r fig_count_norm_DS078, fig.width=8, fig.height=5}
options(repr.plot.width = 6, repr.plot.height = 4)

ggplot(data_properties_norm_true, aes(log10(nCount_norm), nProt, color = cell_treat)) +
  geom_point(alpha = 1, size = 1) +
  facet_grid(~block_buf) +
  # scale_color_manual(values = colors_dark9) +
  theme_bw() +
  labs(title = "Correlation total counts and number of detected barcodes", 
       x = expression("Total counts Log"[10]),
       y = "Detected number of protein barcodes") +
  textsize_small

ggplot(counts_dtbl_norm, aes(log10(counts_target), log10(counts_target_norm_geo), color = cell_treat)) +
  geom_point(alpha = 1, size = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  facet_grid(~block_buf) +
  # scale_color_manual(values = colors_conc, name = "iSYK + iBTK + iNFkB") +
  theme_bw() +
  labs(title = "Correlation original counts and combined counts",
       x = expression("Original counts Log"[10]),
       y = expression("Combined counts Log"[10])) +
  textsize_small
```

***Figure:** Correlation between raw counts and compensated counts.*


### Data wrangling

Calculate the following:

-   Mean + sd for replicates (counts_target + counts_target_norm_geo)

```{r mean_rep_DS078}
# Calculate mean + sd of counts_target and counts_target_norm_geo
mean_dtbl <- counts_dtbl_norm %>%
  dplyr::group_by(description, target_assign) %>%
  summarise(
    counts_sd = sd(counts_target, na.rm = T),
    counts = mean(counts_target, na.rm = T), 
    counts_norm_sd = sd(counts_target_norm_geo, na.rm = T),
    counts_norm = mean(counts_target_norm_geo, na.rm = T)
    )

# Add metadata
mean_dtbl <- mean_dtbl %>%
  left_join(dplyr::select(counts_dtbl,
                   c(barcode_name, DS_Ab_nr, target, DS_barcode_nr, DS_barcode_seq, target_nospace, target_assign, target_assign_unique, cell_line, stimulus, stim_clean, stim_stain, block_buf, cell_treat, staining, description))) %>%
  distinct()
```



### Save data

Finally, we will save the datasets as .csv file

-   IDseq_data_sample.csv: Dataset of counts per protein per sample 

-   IDseq_data_condition.csv: Dataset of counts per protein per condition (mean of replicates)

```{r save_DS078}
# Dataset of counts per protein per sample 
meta_cols_well_DS078 <- c("plate_well", "plate", "well", "experiment", "cell_line", "stimulus", "stim_clean", "stim_stain", "block_buf", "cell_treat", "replicate", "description", "description_rep",
                          "target_nospace", "target_assign", "target_assign_unique", "barcode_name", "DS_Ab_nr", 
                          "counts", "counts_target", "true_scaling_factor_geo", "counts_target_norm_geo")
data_well_all <- counts_dtbl_norm %>% 
  arrange(plate_well, target_assign) %>%
  dplyr::select(meta_cols_well_DS078) %>% # reorder columns for a more logical order
  dplyr::rename(Ab_BC_nr = barcode_name, 
                scaling_factor = true_scaling_factor_geo, 
                counts_target_norm = counts_target_norm_geo)

write.csv(data_well_all, file = "output/DS078_Blocking1/IDseq_ann/IDseq_data_sample.csv", row.names = F)

# Dataset of mean counts per protein per condition (replicates combined)
meta_cols_mean_DS078 <- c("cell_line", "stimulus", "stim_clean", "stim_stain", "block_buf", "cell_treat", "description",
                          "target_nospace", "target_assign", "target_assign_unique", "barcode_name", "DS_Ab_nr", 
                          "counts", "counts_sd", "counts_norm", "counts_norm_sd")
data_mean_all <- mean_dtbl %>%
  dplyr::select(meta_cols_mean_DS078) %>% # reorder cols to more logical order
  dplyr::rename(Ab_BC_nr = barcode_name)

write.csv(data_mean_all, file = "output/DS078_Blocking1/IDseq_ann/IDseq_data_condition.csv", row.names = F)
```

```{r clear_DS078}
rm(list = ls(pattern = "DS078"))
rm(list = ls(pattern = "Plate"))
rm(list = ls(pattern = "comp"))
gc()
```



## DS091: ImmunostainingBio

### Raw data

```{r data_DS091}
# Load sparse matrix format all data
counts_PlateA <- read_count_output("data/DS091_ImmunostainingBio/counts/PlateA/featurecounts", name = "featurecounts")
meta_wellID <- read_csv("data/DS091_ImmunostainingBio/config/DynSign.091_wellBC_metadata_seq.csv")

meta_Abs <- read_excel("data/DS091_ImmunostainingBio/config/BulkIDseq_Ab_info_seq_analysis_202212.xlsx")

# Reshape counts in data table format
counts_dtbl <- data.frame(counts_PlateA) %>% 
  mutate(barcode_name = rownames(counts_PlateA)) %>% #barcode_name #Barcodename
  dplyr::select(barcode_name, everything()) %>%
  gather("well_BC_seq", "counts", 2:c(ncol(counts_PlateA)+1)) %>%
  dplyr::filter(counts >= 1) %>% # Remove undetected proteins counts
  left_join(meta_Abs) %>%
  unique() %>%
  # dplyr::filter(!is.na(target_nospace)) %>% # Remove barcodes not in Ab metadata (not conjugated/added in panel)
  left_join(subset(meta_wellID, plate == "A")) %>% # Add sample metadata
  dplyr::filter(!is.na(experiment)) %>%
  arrange(plate, well)

counts_dtbl <- counts_dtbl %>%
  mutate(plate_well = paste0(plate, "_", well))

# Calculate per-sample properties
data_properties <- counts_dtbl %>%
  unique()%>%
  group_by(plate, well_BC_seq) %>%
  summarize(nCount = sum(counts), 
            nProt = n()) %>%
  left_join(meta_wellID) # Add sample metadata 
```

***Table**: Unfiltered data-table properties loaded into R*

|                           | nCols (#samples)        | nRows (\# barcodes)     |
|------------------|---------------------------|---------------------------|
| Plate A         | `r ncol(counts_PlateA)` | `r nrow(counts_PlateA)` |

The full panel of mapped barcodes and 96 wells are present in the data table. First have a look at the total counts and number of detected protein barcodes per sample: There are a lot more proteins detected than put in the panel, due to barcode contamination and cross-ver between wells. The nostain samples have relatively low counts and proteins detected.

```{r fig_count_prot_DS091, fig.width=8, fig.height=5}
options(repr.plot.width = 6, repr.plot.height = 4)

ggplot(data_properties, aes(log10(nCount), nProt, color = as.character(staining_conc))) +
  geom_point(alpha = 1, size = 1) +
  facet_wrap(~stimulus) + # stimulus, staining_cells, staining_conc, staining_volume
  scale_color_manual(values = colors_blue9[2:7], name = "Staining conc \n(ug/mL)") +
  # scale_color_binned(type = "viridis", name = "Staining conc (ug/mL)") +
  theme_bw() +
  labs(title = "Correlation total counts and number of detected barcodes", 
       x = expression("Total counts Log"[10]),
       y = "Detected number of protein barcodes") +
  textsize_small
```

***Figure:** Correlation between total counts and number of detected protein barcodes. We observe many more detected proteins than present in the staining.*

As we have a 96-wells plate, we can also have a look at total counts per plate detected:

```{r fig_plate_totalUMI_DS091, fig.width=8, fig.height=4}
options(repr.plot.width = 10, repr.plot.height = 6)

rng <- range(log10(data_properties$nCount))

plot <- raw_grid(data = log10(data_properties$nCount), well = data_properties$well, plate_id = data_properties$plate, plate = 96, size = 6) +
  scale_fill_viridis(expression("Total counts Log"[10]), limits = c(rng[1], rng[2])) +
  theme_bw()

print(plot)

```

***Figure:** Plate overviews of total counts.*


### Barcode contamination comp

Several barcodes are contaminated with \<2.5% of other barcodes (all barcodes were sequenced individually to determine the contamination). As this can skew the results, especially for lowly present barcodes that are contaminated with highly present barcodes, we will compensate for this.

```{r comp_apply_DS091}
comp_matrix <- read_csv("data/DS091_ImmunostainingBio/BC_compensation_matrix.csv")

data_comp <- counts_dtbl %>%
  # dplyr::filter(well == "A01") %>%
  dplyr::select(plate, well, barcode_name, counts) %>%
  left_join(comp_matrix, by = join_by(barcode_name == BC_to_correct)) #, multiple = "all"

# Get counts of OG BC
data_comp_OG <- data_comp %>%
  # dplyr::filter(well == "A01") %>%
  dplyr::filter(barcode_name == OG_BC) %>%
  dplyr::select(plate, well, OG_BC, counts) %>%
  dplyr::rename(counts_OG_BC = counts)

# Calculate false reads
data_false <- left_join(data_comp, data_comp_OG) %>% 
  mutate(false_counts = (counts_OG_BC / percent_OG_BC) * percent_BC_to_correct) %>%
  dplyr::filter(barcode_name != OG_BC) %>%
  group_by(plate, well, barcode_name, counts) %>%
  summarize(total_false_counts = sum(false_counts, na.rm = TRUE))

counts_dtbl <- counts_dtbl %>%
  # dplyr::filter(well == "A01") %>%
  left_join(data_false) %>% 
  mutate(total_false_counts = replace_na(total_false_counts, 0), 
         true_counts = as.integer(counts - total_false_counts), 
         percent_true = (true_counts / counts) * 100) %>%
  dplyr::filter(!is.na(target_nospace)) # Remove barcodes not in Ab metadata (not conjugated/added in panel)

compensation <- counts_dtbl %>%
  dplyr::select(plate, well, barcode_name, counts, total_false_counts, true_counts, percent_true, target_nospace)

# Calculate per-sample properties
data_properties_true <- counts_dtbl %>%
  unique()%>%
  group_by(plate, well_BC_seq) %>%
  summarize(nCount = sum(true_counts), 
            nProt = n()) %>%
  left_join(meta_wellID)
```

```{r fig_count_raw_comp_DS091, fig.width=8, fig.height=5}
options(repr.plot.width = 6, repr.plot.height = 4)

ggplot(counts_dtbl, aes(log10(counts), log10(true_counts), color = as.character(staining_conc))) +
  geom_point(alpha = 1, size = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  facet_wrap(~stimulus) +
  scale_color_manual(values = colors_blue9[4:9], name = "Staining conc \n(ug/mL)") +
  theme_bw() +
  labs(title = "Correlation raw counts and compensated counts",
       x = expression("Raw counts Log"[10]),
       y = expression("Compensated counts Log"[10])) +
  textsize_small

ggplot(data_properties_true, aes(log10(nCount), nProt, color = as.character(staining_conc))) +
  geom_point(alpha = 1, size = 1) +
  facet_wrap(~stimulus) + # stimulus, staining_cells, staining_conc, staining_volume
  scale_color_manual(values = colors_blue9[2:7], name = "Staining conc \n(ug/mL)") +
  # scale_color_binned(type = "viridis", name = "Staining conc (ug/mL)") +
  theme_bw() +
  labs(title = "Correlation total counts and number of detected barcodes", 
       x = expression("Total counts Log"[10]),
       y = "Detected number of protein barcodes") +
  textsize_small
```

***Figure:** Correlation between raw counts and compensated counts, and between compensated counts and detected number of barcodes. After compensation, all samples (except nostain samples) have 107 detected proteins (the number of Ab in the panel).*


### Count normalisation

To remove variation between replicates, the data is normalised per target protein with a geometric mean scaling factor: First, the geometric mean of each target is calculated separately for each cell line, and then the median of the protein scaling factors is calculated separately for each sample.

```{r normalisation_DS091}
scaling_factors <- counts_dtbl %>%
  dplyr::filter(cell_line %in% c("HBL1", "DAUDI")) %>%
  # filter() %>% #target_nospace == "GAPDH" # modification == "no"
  ungroup() %>%
  group_by(cell_line, target_nospace) %>%
  mutate(scaling_factor = counts / median(counts),
         scaling_factor_geo = counts / exp(mean(log((counts)))), 
         true_scaling_factor = true_counts / median(true_counts),
         true_scaling_factor_geo = true_counts / exp(mean(log((true_counts))))) %>%
  ungroup() %>%
  group_by(cell_line, description_sample, description_staining) %>%
  summarize(scaling_factor = median(scaling_factor), .groups = "keep",
            scaling_factor_geo = median(scaling_factor_geo), 
            true_scaling_factor = median(true_scaling_factor),
            true_scaling_factor_geo = median(true_scaling_factor_geo, na.rm = TRUE))

counts_dtbl_norm <- counts_dtbl %>%
  dplyr::filter(cell_line %in% c("HBL1", "DAUDI")) %>%
  left_join(scaling_factors) %>%
  ungroup() %>%
  mutate(counts_norm = counts / scaling_factor,
         counts_norm_geo = counts / scaling_factor_geo, 
         true_counts_norm = true_counts / true_scaling_factor, 
         true_counts_norm_geo = true_counts / true_scaling_factor_geo)

counts_dtbl_norm <- dplyr::filter(counts_dtbl_norm, !is.na(target_nospace)) %>%
  mutate(stimulus = factor(stimulus, levels = c("PBS", "anti-Ig", "H2O2")))

# Calculate per-sample properties
data_properties_norm <- counts_dtbl_norm %>%
  unique()%>%
  group_by(well_BC_seq) %>%
  summarize(nCount_norm = sum(counts_norm_geo), 
            nProt = n()) %>%
  left_join(subset(meta_wellID, plate == "A")) %>% # Add sample metadata 
  mutate(stimulus = fct_relevel(stimulus, "PBS", "anti-Ig", "H2O2"))

data_properties_norm_true <- counts_dtbl_norm %>%
  unique()%>%
  group_by(well_BC_seq) %>%
  summarize(nCount_norm = sum(true_counts_norm_geo), 
            nProt = n()) %>%
  left_join(subset(meta_wellID, plate == "A")) %>% # Add sample metadata 
  mutate(stimulus = fct_relevel(stimulus, "PBS", "anti-Ig", "H2O2"))
```

```{r fig_count_norm_DS091, fig.width=8, fig.height=5}
options(repr.plot.width = 6, repr.plot.height = 4)

ggplot(counts_dtbl_norm, aes(log10(counts), log10(counts_norm_geo), color = as.character(staining_conc))) +
  geom_point(alpha = 1, size = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  facet_wrap(~stimulus) +
  scale_color_manual(values = colors_blue9[4:9], name = "Staining conc \n(ug/mL)") +
  theme_bw() +
  labs(title = "Correlation raw counts and normalised counts",
       x = expression("Raw counts Log"[10]),
       y = expression("Normalised counts Log"[10])) +
  textsize_small

ggplot(data_properties_norm, aes(log10(nCount_norm), nProt, color = as.character(staining_conc))) +
  geom_point(alpha = 1, size = 1) +
  facet_wrap(~stimulus) + # stimulus, staining_cells, staining_conc, staining_volume
  scale_color_manual(values = colors_blue9[2:7], name = "Staining conc \n(ug/mL)") +
  # scale_color_binned(type = "viridis", name = "Staining conc (ug/mL)") +
  theme_bw() +
  labs(title = "Correlation normalised counts and number of detected barcodes", 
       x = expression("Normalised counts Log"[10]),
       y = "Detected number of protein barcodes") +
  textsize_small
```

***Figure:** Correlation between raw counts and compensated counts.*

```{r normalisation_bio_DS091}
scaling_factors_bio <- counts_dtbl %>%
  dplyr::filter(staining_conc_text == "0.125 ug/mL") %>%
  # filter() %>% #target_nospace == "GAPDH" # modification == "no"
  ungroup() %>%
  group_by(cell_line, target_nospace) %>%
  mutate(scaling_factor = counts / median(counts),
         scaling_factor_geo = counts / exp(mean(log((counts)))), 
         true_scaling_factor = true_counts / median(true_counts),
         true_scaling_factor_geo = true_counts / exp(mean(log((true_counts))))) %>%
  ungroup() %>%
  group_by(cell_line, description_sample, description_staining) %>%
  summarize(scaling_factor = median(scaling_factor), .groups = "keep",
            scaling_factor_geo = median(scaling_factor_geo), 
            true_scaling_factor = median(true_scaling_factor),
            true_scaling_factor_geo = median(true_scaling_factor_geo, na.rm = TRUE))

counts_dtbl_bio <- counts_dtbl %>%
  dplyr::filter(staining_conc_text == "0.125 ug/mL") %>%
  left_join(scaling_factors) %>%
  ungroup() %>%
  mutate(counts_norm = counts / scaling_factor,
         counts_norm_geo = counts / scaling_factor_geo, 
         true_counts_norm = true_counts / true_scaling_factor, 
         true_counts_norm_geo = true_counts / true_scaling_factor_geo)

counts_dtbl_bio <- dplyr::filter(counts_dtbl_bio, !is.na(target_nospace)) %>%
  mutate(stimulus = factor(stimulus, levels = c("PBS", "anti-Ig", "H2O2")))

ggplot(counts_dtbl_bio, aes(log10(counts), log10(counts_norm_geo), color = stim_clean)) +
  geom_point(alpha = 1, size = 0.8) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + 
  facet_wrap(~cell_line) +
  scale_color_manual(values = colors_stim_clean, name = "Stimulus") +
  theme_bw() +
  labs(title = "Correlation raw counts and normalised counts",
       x = expression("Raw counts Log"[10]),
       y = expression("Normalised counts Log"[10])) +
  textsize_small
```


### Data wrangling

Calculate the following:

-   Mean + sd for replicates (true_counts + true_counts_norm_geo)

```{r mean_rep_DS091}
# Calculate mean + sd of true_counts and true_counts_norm_geo
# Dataset with all samples
mean_dtbl_norm <- counts_dtbl_norm %>%
  dplyr::group_by(description_sample, description_staining, target_nospace) %>%
  summarise(
    counts_sd = sd(true_counts, na.rm = T),
    counts = mean(true_counts, na.rm = T), 
    counts_norm_sd = sd(true_counts_norm_geo, na.rm = T),
    counts_norm = mean(true_counts_norm_geo, na.rm = T)
    )

# Add metadata
mean_dtbl_norm <- mean_dtbl_norm %>%
  left_join(dplyr::select(counts_dtbl,
                   c(barcode_name, DS_Ab_nr, target, modification, epitope, marker_cell, marker_pathway, marker_pathway_detail, isotype, DS_barcode_nr, DS_barcode_seq, target_nospace, cell_line, stimulus, stim_clean, staining_cells, staining_conc, staining_conc_text, staining_volume, description_sample, description_staining))) %>%
  distinct()

# Dataset with bio samples
mean_dtbl_bio <- counts_dtbl_bio %>%
  dplyr::group_by(description_sample, description_staining, target_nospace) %>%
  summarise(
    counts_sd = sd(true_counts, na.rm = T),
    counts = mean(true_counts, na.rm = T), 
    counts_norm_sd = sd(true_counts_norm_geo, na.rm = T),
    counts_norm = mean(true_counts_norm_geo, na.rm = T)
    )

# Add metadata
mean_dtbl_bio <- mean_dtbl_bio %>%
  left_join(dplyr::select(counts_dtbl,
                   c(barcode_name, DS_Ab_nr, target, modification, epitope, marker_cell, marker_pathway, marker_pathway_detail, isotype, DS_barcode_nr, DS_barcode_seq, target_nospace, cell_line, stimulus, stim_clean, staining_cells, staining_conc, staining_conc_text, staining_volume, description_sample, description_staining))) %>%
  distinct()
```



### Save data

Finally, we will save the datasets as .csv file

-   IDseq_data_sample.csv: Dataset of counts per protein per sample (all samples)

-   bio_IDseq_data_sample.csv: Dataset of counts per protein per sample (bio samples)

-   IDseq_data_condition.csv: Dataset of counts per protein per condition (mean of replicates) (all samples)

-   bio_IDseq_data_condition.csv: Dataset of counts per protein per condition (mean of replicates) (bio samples)

```{r save_DS091}
# Dataset of counts per protein per sample 
meta_cols_well_DS091 <- c("plate_well", "plate", "well", "experiment", "cell_line", "stimulus", "stim_clean", "replicate", "staining", "staining_cells", "staining_conc", "staining_conc_text", "staining_volume", "description_sample", "description_sample_rep", "description_staining",
                    "target_nospace", "target", "modification", "marker_cell", "marker_pathway", "marker_pathway_detail", "barcode_name", "DS_Ab_nr", 
                    "true_counts", "true_scaling_factor_geo", "true_counts_norm_geo")
data_well_all <- counts_dtbl_norm %>% 
  arrange(plate_well, target_nospace) %>%
  dplyr::select(meta_cols_well_DS091) %>% # reorder columns for a more logical order
  dplyr::rename(marker_path = marker_pathway, # rename some columns for clarity
                marker_path_detail = marker_pathway_detail, 
                Ab_BC_nr = barcode_name, 
                counts = true_counts, 
                scaling_factor = true_scaling_factor_geo, 
                counts_norm = true_counts_norm_geo)
write.csv(data_well_all, file = "output/DS091_ImmunostainingBio/IDseq_ann/IDseq_data_sample.csv", row.names = F)

data_well_bio <- counts_dtbl_bio %>% 
  arrange(plate_well, target_nospace) %>%
  dplyr::select(meta_cols_well_DS091) %>% # reorder columns for a more logical order
  dplyr::rename(marker_path = marker_pathway, # rename some columns for clarity
                marker_path_detail = marker_pathway_detail, 
                Ab_BC_nr = barcode_name, 
                counts = true_counts, 
                scaling_factor = true_scaling_factor_geo, 
                counts_norm = true_counts_norm_geo)
write.csv(data_well_bio, file = "output/DS091_ImmunostainingBio/IDseq_ann/bio_IDseq_data_sample.csv", row.names = F)

# Dataset of mean counts per protein per condition (replicates combined)
meta_cols_mean_DS091 <- c("cell_line", "stimulus", "stim_clean", "staining_cells", "staining_conc", "staining_conc_text", "staining_volume", "description_sample", "description_staining", 
                    "target_nospace", "target", "modification", "marker_cell", "marker_pathway", "marker_pathway_detail", "barcode_name", "DS_Ab_nr", 
                    "counts", "counts_sd", "counts_norm", "counts_norm_sd")
data_mean_all <- mean_dtbl_norm %>%
  dplyr::select(meta_cols_mean_DS091) %>% # reorder cols to more logical order
  dplyr::rename(marker_path = marker_pathway, # rename some columns for clarity
                marker_path_detail = marker_pathway_detail, 
                Ab_BC_nr = barcode_name)
write.csv(data_mean_all, file = "output/DS091_ImmunostainingBio/IDseq_ann/IDseq_data_condition.csv", row.names = F)

data_mean_bio <- mean_dtbl_bio %>%
  dplyr::select(meta_cols_mean_DS091) %>% # reorder cols to more logical order
  dplyr::rename(marker_path = marker_pathway, # rename some columns for clarity
                marker_path_detail = marker_pathway_detail, 
                Ab_BC_nr = barcode_name)
write.csv(data_mean_bio, file = "output/DS091_ImmunostainingBio/IDseq_ann/bio_IDseq_data_condition.csv", row.names = F)
```

```{r clear_DS091}
rm(list = ls(pattern = "DS091"))
rm(list = ls(pattern = "Plate"))
rm(list = ls(pattern = "comp"))
gc()
```
